Semantic browsing of sound databases \\ without keywords

### 2016-06-08 responseToReviewers.tex

\section{Responses to Reviewer 2}

\begin{enumerate}

\item \emph{If the author replicate the same experiment for example using a music collection or a collection of bird songs and the findings were consistent 
this would justify the old title (and make the paper much stronger)}

This is an excellent suggestion. We consider an new dataset of sounds produced by musical instruments. As shown on Section 6 of the updated paper, the findings are consistent with the ones obtained using the urban sound dataset.

We thus choose to stay with the original title.

@@

Comparing to an optimized SOM display would have been interested, along with the use of other features. That said, we decided to dedicate the complementary experiments to the use of an alternate dataset.

\item \emph{some thing 
as simple as showing the word associated with a category when navigating would probably help. This would be an easy configuration to test and I am surprised the authors [not]
have done so. I suspect that the speed and efficiency would improve.
}

We agree that this configuration would have helped us gaining a more precise understanding of the results. However, we chose to focus our study on completely non verbal displays as stated in the introduction.

Even if this complementary study would have increase the level of knowledge, we again decided to focus the complementary experiments to the use of an alternative dataset.

### 2016-06-08 lafaySsf2016.tex
revision version

\usepackage{color,url}
\newcommand{\gl}[1]{\textcolor{red}{gl: #1}}
\newcommand{\mr}[1]{\textcolor{green}{Mathias : #1}}
\newcommand{\ml}[2][]{\textcolor{blue}{#1 #2}}
\newcommand{\nm}[1]{\textcolor{magenta}{nm: #1}}
%\renewcommand{\baselinestretch}{1.5}

@@

We present and study the efficiency of two sound presentation strategies that organize sounds spatially so as to reflect an underlying semantic hierarchy. For the sake of comparison, we also consider a display whose spatial organization is based on acoustic cues. Those three displays are evaluated in terms of search speed in a crowdsourcing experiment using two different corpora: the first is composed of environmental sounds from urban environments and the second of sounds produced by musical instruments. Coherent results achieved by considering the two different corpora demonstrate the usefulness of using an implicit semantic organization to display sounds, both in terms of search speed and of learning efficiency. 
\end{abstract}

@@

Most media asset management systems are based on keyword-driven queries. The user enters a word which best characterizes the desired item, and the interface presents him with items related to this word. The effectiveness of this principle is primarily based on the typological structure and nomenclature of the database. However, for sound databases, several issues arise:

\begin{enumerate}
\item Sounds, as many others things, can be described in many ways. Environmental sounds in particular may be designated by their
source (a car door),  the action of that source (the slamming of a car door) or their environment (slamming a car door in a garage) \cite{houix2012lexical, niessen2010categories, brown2011towards}. Designing an effective keyword-based search system requires an accurate description of each sound, which has to be tunable to the sound representation of each user.
\item Pre-defined verbal descriptions of the sounds made available to the users may potentially bias their browsing and final selection.
\item Localization of the query interface is made difficult as the translation of some words referring to qualitative aspects of the sound, such as its ambiance, is notoriously ambiguous and subject to cultural specificities.
\item Unless considerable time and resources are invested into developing a multilingual interface, any system based on verbal descriptions can only be used with reduced performance by non-native speakers of the chosen language.
\end{enumerate}
To circumvent those issues, not relying on keywords is desirable. However, conveying semantics remains a necessity for ease of browsing, which raises the question of what alternate means, other than written language, can be used to that end.

@@

Similarly, the second corpus also has a semantic organization: high and mid level classes categorize musical instruments according to the means by which sound is produced, while leaf level classes are related to specific instruments (\textit{string} $>$ \textit{plucked-string} $>$ \textit{viola}; \gl{\textit{brass} $>$ \textit{ordinario} $>$ \textit{trumpet}}).

@@

Among the several alternatives that exists, MFCCs features are considered as they are widely used in the audio processing community. MDS is selected for 2D projection as it has lower sensitivity to outliers than the PCA and has the advantage of not relying on any parameter setting unlike the SOM approach.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/music_mds.png} 
\end{center}
\caption{\label{figXP3music} Acoustical Display for the musical sound dataset (AD). Colors and legend are removed during the experiment. \gl{remove ?}}
\end{figure}

@@

\mr{À enlever si vous êtes d'accord, c'est répété juste en-dessous : The PSD interface is introduced in order to study if a progressive top-down display of the hierarchical structure helps the user explore and understand the dataset hierarchy.}

@@

During the experiment, two visual cues are shown to the subject:
\begin{itemize}
\item \textit{The search state}: which can be "pause" if the subject is not currently looking for a target sound (at the beginning of the experiment, or between two target sounds) or  "in progress"  if they are.
\item \textit{Remaining target sounds}: the number of target sounds that remain to be found.
\end{itemize}

@@

A crowdsourcing approach is adopted. The experiment is designed to be performed using the \textit{Chrome} web browser. The link to the experiment has been sent to three  mailing lists, namely \textit{Music-Ir}, \textit{Auditory} and an internal IRCAM mailing list. Subjects are only allowed to perform the experiment once, and on one interface only. Data were automatically collected server-side at the end of the experiments. Subjects were asked to use headphones. All sounds of the dataset are normalized to the same root mean square (RMS) level.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/urbanMusic1.eps} 
\end{center}
\caption{\label{urbanMusic} \gl{Average and standard deviation for (a,d) the total search durations, (b,e) the number of heard sounds and (c,f) the number of unique heard sounds; considering the urban dataset (a,b,c) or the music dataset (d,e,f)}}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/analyse_urban2.eps} 
\end{center}
\caption{\label{fig2urban} \gl{evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the urban dataset. Lines are linear regression fits.}}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/analyse_music2.eps} 
\end{center}
\caption{\label{fig2Music} \gl{evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the music dataset. Lines are linear regression fits.}}
\end{figure}

@@

According to those results, a spatial display following a hierarchical organization of the dataset based on semantic values (PSD and FSD) is more efficient than an organization based on acoustic descriptors (AD) as it allows the users to quickly apprehend the dataset range and thus to specify their search. Furthermore the FSD interface allows the users to retrieve the 13 target sounds by listening to a smaller amount of sounds, than the AD. However, those two effects are significantly compromised when users have to parse the entire hierarchy to reach the first target sound, as in the PSD display. It seems that enforcing an explicit exploration of the hierarchy disturbs or confuses the user instead of allowing him to learn the semantically motivated spatial organization of the classes.

@@

In this paper, two displays allowing users to explore a semantically organized sound dataset without written textual help are presented. The interfaces distribute sounds represented by circles on a 2D space, following an underlying hierarchical semantic organization. Those two semantic displays are assessed and compared to a third display in which the spatial configuration depends upon acoustic features. The tests consists in data retrieval tasks done using two different corpora. 
The Full Semantic Display (FSD), that allows users to directly visualize the leaf classes of the semantic hierarchical structure, proves to be the most effective interface for the task.

@@

In the datasets considered in this study, there are as many leaf classes as sounds in the dataset, which would be unrealistic for large datasets. The organization can easily be adapted by considering the leaf classes as \textit{collections} of semantically similar sound samples. Thus, two sounds of \textit{male-yelling} would be grouped into a single leaf class with the tag \textit{male-yelling}. The leaf class would then also have a \textit{prototype sound} being the most representative item of the different \textit{male-yelling} sounds belonging to the leaf class.

>>>>

\usepackage{color,url}
\newcommand{\gl}[1]{\textcolor{red}{gl: #1}}
\newcommand{\mr}[1]{\textcolor{green}{Mathias : #1}}
\newcommand{\ml}[2][]{\textcolor{blue}{#1 #2}}
\newcommand{\nm}[1]{\textcolor{magenta}{nm: #1}}
%\renewcommand{\baselinestretch}{2}

@@

We present and study the efficiency of two sound presentation strategies that organize sounds spatially so as to reflect an underlying semantic hierarchy. For the sake of comparison, we also consider a display whose spatial organization is only based on acoustic cues. Those three displays are evaluated in terms of search speed in a crowdsourcing experiment using two different corpora: the first is composed of environmental sounds from urban environments and the second of sounds produced by musical instruments. Coherent results achieved by considering the two different corpora demonstrate the usefulness of using an implicit semantic organization to display sounds, both in terms of search speed and of learning efficiency. 
\end{abstract}

@@

Thus, most media asset management systems are based on keyword-driven queries. The user enters a word which best characterizes the desired item, and the interface presents him with items related to this word. The effectiveness of this principle is primarily based on the typological structure and nomenclature of the database. However, for sound databases, several issues arise:

\begin{enumerate}
\item Sounds, as many others things, can be described in many ways. Environmental sounds in particular may be designated by their
source (a car door),  the action of that source (the slamming of a car door) or their environment (slamming a car door in a garage) \cite{houix2012lexical, niessen2010categories, brown2011towards}. Designing an effective keyword-based search system requires an accurate description of each sound, which has to be suited to the sound representation of the user to be effective.
\item Pre-defined verbal descriptions of the sounds made available to the users may potentially bias their browsing and final selection.
\item Localization of the query interface is made difficult as the translation of some words referring to qualitative aspects of the sound, such as its ambiance, is notoriously ambiguous and subject to cultural specificities.
\item Unless considerable time and resources are invested into developing a multilingual interface, any system based on verbal descriptions can only be used with reduced performance by non-native speakers of the chosen language.
\end{enumerate}
To circumvent those issues, not relying on keywords is desirable. However, conveying semantics remains a necessity for ease of browsing, which raises the question of what alternate means, other than written language, can be used to that end.

@@

Similarly, the second corpus also has a semantic organization: high and mid level classes categorize musical instruments according to the means by which sound is produced, while leaf level classes are related to specific instruments (\textit{string} $>$ \textit{plucked-string} $>$ \textit{viola}; \textit{brass} $>$ \textit{ordinario} $>$ \textit{trumpet}).

@@

Among the several alternatives that exists, MFCCs features are chosen as they are widely used in the audio processing community. MDS is selected for 2D projection as it has a lower sensitivity to outliers than the PCA and do not relyi on any parameter optimization, unlike the SOM approach.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/music_mds.png} 
\end{center}
\caption{\label{figXP3music} Acoustical Display for the musical sound dataset (AD). Colors and legend are removed during the experiment.}
\end{figure}

@@

%\mr{À enlever si vous êtes d'accord, c'est répété juste en-dessous : The PSD interface is introduced in order to study if a progressive top-down display of the hierarchical structure helps the user explore and understand the dataset hierarchy.}

@@

%During the experiment, two visual cues are shown to the subject:
%\begin{itemize}
%\item \textit{The search state}: which can be "pause" if the subject is not currently looking for a target sound (at the beginning of the experiment, or between two target sounds) or  "in progress"  if they are.
%\item \textit{Remaining target sounds}: the number of target sounds that remain to be found.
%\end{itemize}

@@

A crowdsourcing approach is adopted. 
%The experiment is designed to be performed using the \textit{Chrome} web browser. The link to the experiment has been sent to three  mailing lists, namely \textit{Music-Ir}, \textit{Auditory} and an internal IRCAM mailing list. 
Subjects are recruited using research mailing list and are only allowed to perform the experiment once, and on one interface only. % Data were automatically collected server-side at the end of the experiments. Subjects were asked to use headphones. 
All sounds of the dataset are normalized to the same root mean square (RMS) level.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/urbanMusic1.eps} 
\end{center}
\caption{\label{urbanMusic} Average and standard deviation for (a,d) the total search durations, (b,e) the number of heard sounds and (c,f) the number of unique heard sounds; considering respectively the urban dataset (a,b,c) and the music dataset (d,e,f).}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/analyse_urban2.eps} 
\end{center}
\caption{\label{fig2urban} Evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the urban dataset. Lines are linear regression fits.}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/analyse_music2.eps} 
\end{center}
\caption{\label{fig2Music} Evolution of (a) the average search durations, (b) the average numbers of heard sounds, and (c) the average numbers of unique heard sounds at each target sound search, considering the music dataset. Lines are linear regression fits.}
\end{figure}

@@

According to those results, a spatial display following a hierarchical organization of the dataset based on semantic values (PSD and FSD) is more efficient than an organization based on acoustic descriptors (AD) as it allows the users to quickly apprehend the dataset range and thus to specify their search. Furthermore the FSD interface allows the users to retrieve the 13 target sounds by listening to a smaller amount of sounds than the AD. However, those two effects are significantly compromised when users have to parse the entire hierarchy to reach the first target sound, as in the PSD display. It seems that enforcing an explicit exploration of the hierarchy disturbs or confuses the user instead of allowing him to learn the semantically motivated spatial organization of the classes.

@@

In this paper, two displays allowing users to explore a semantically organized sound dataset without written textual help are presented. The interfaces distribute sounds represented by circles on a 2D space, following an underlying hierarchical semantic organization. Those two semantic displays are assessed and compared to a third display in which the spatial configuration depends upon acoustic features. The tests consists in data retrieval tasks done using two different corpora. 
The Full Semantic Display (FSD), that allows users to directly visualize the leaf classes of the semantic hierarchical structure, proves to be the most effective interface for the task. This display has been integrated as a sound selector into a acoustic scene simulator that benefit from the advantages of such an approach: no need of localization and reduction of the potential bias of using description tags\footnote{\url{http://soundthings.org/research/urbanSoundScape/XP2014/simScene}}.

@@

In the datasets considered in this study, there are as many leaf classes as sounds in the dataset, which would be unrealistic for larger datasets. In that case, the organization can easily be adapted by considering the leaf classes as \textit{collections} of semantically similar sound samples. Thus, two sounds of \textit{male-yelling} would be grouped into a single leaf class with the tag \textit{male-yelling}. The leaf class would then also have a \textit{prototype sound} being the most representative item of the different \textit{male-yelling} sounds belonging to the leaf class.
